A Compound Decision Approach to Covariance Matrix Estimation

### 2022-03-08 response_reviews.tex

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[round]{natbib}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{comment}
\usepackage{lscape}

\usepackage{xr}
\externaldocument{msg}

@@

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\def\bs{\boldsymbol}
\def\bb{\mathbb}

\title{Response to reviews of BIOM2021236M}
\author{}
\date{}

\begin{document}
\maketitle

\section{Response to Reviewer 2}

Thank you for your very careful second reading and the excellent comments. We respond to each point below.

\begin{enumerate}
\item \emph{Population and sample eigenvectors: A key thesis of the revised manuscript that is emphasized multiple times (even in the abstract) is that the new method outperforms optimal rotationally invariant methods in situations where sample eigenvectors are far from population eigenvectors. I think this claim is not suffciiently supported \ldots I think either more compelling evidence should be provided, or the claim can be downemphasized and perhaps mentioned once in the manuscript as a possible conjecture.}

  This makes sense. We have downplayed this point and merely mention it as a conjecture; \textbf{see the last paragraph of Section 4}.

\item \emph{Gene networks in Figures 4 and 5: I found these networks diffciult to interpret. It would greatly help if the presentation of these figures could be improved. Perhaps one could present less methods, but superimpose resulting edges onto a joint set of nodes?}

  Thanks for the great suggestion, we have adopted your approach; \textbf{see Figures 3 and 4 and a discussion of the figures at the end of Section 5}.

\item \emph{counts per million mapped reads: My understanding is that this method of normalizing counts is not recommended for RNA-Seq experiments. For example, DESeq2 uses another method of normalization. Why was this method chosen? Do the results change with other normalization methods? }

  There are indeed a number of normalization methods, but most of the work in this area has been for identifying differentially expressed genes. It is not always clear what normalization to use for more complex analyses, such as the coexpression network study in this paper. In particular, the most recent version of the edgeR user's guide states in Section 2.16 that

  ``Inputing RNA-seq counts to clustering or heatmap routines designed for microarray data is not straight-forward, and the best way to do this is still a matter of research. To draw a heatmap of individual RNA-seq samples, we suggest using moderated log-counts-per-million.''

  We have followed this CPM approach in our analyses. CPM normalization for coexpression has also been shown to have good performance \citep{johnson2022robust}.

\item \emph{Could the authors provide a bit more detail on splitting into training and testing set? Is time ignored when splitting the samples? Also is the sample covariance (that first centers observations; as in the simulation) used for both test and training samples?}

  We do ignore time, for the sake of sample size. This was also done in the original Saul et al. (2017) paper. We also state that we use the sample covariance matrix, which first centers the observations; \textbf{see the second-to-last paragraph of Section 5}.

\item \emph{OracMSG: (method defined on Page 16): Why do support points need to be generated here? Since all the ($\sigma_j$, $\sigma_k$, $r_{jk}$) are known for the oracle method, I would think that the oracle rule can be computed directly without any need for optimization.}

  We performed clustering even though the true parameters are all known in order to reduce the number of support points to ease the computational burden. When $p$ is large the number of true support points could be very large, which dramatically increased the computation time. We have now mentioned this; \textbf{see Section 4.2}.

\item \emph{Some notation on indexing the population/sample covariance matrix should be made more explicit in the manuscript. In particular, it appears that throughout the text there is the convention on single/double-indexing that $\sigma^2_j = \sigma_{jj}$ and $s^2_j = s_{jj}$ that confused me in light of e.g. (2). (Similarly for $r_{jk}$.)}

  We have now clarified that $\sigma_{jj} = \sigma_j^2$; \textbf{see the first paragraph of Section 2.1}.

\item \emph{Page 17, ``The corrected version consistently outperformed the uncorrected version.'': I think this should hold provably (not only just empirically) by a convex analysis projection argument.}

  Thanks for this insight! However, we do not exactly project our matrix onto the space of positive semi-definite matrices because we want positive-definiteness. Our minimum eigenvalue is $c_\alpha$, as we describe in (14), and we haven't yet thought through how this would affect the argument.

\item \emph{Page 3, ``The article is organized as follows.'', something is off with the numbering of Sections in this paragraph. }

  Thank you for pointing this out, we have fixed the numbering.

\item \emph{Page 4, ``incentived'' typo }

  Fixed, thanks.

\item \emph{Page 4, ``about index i ... different index j'': Maybe phrase this in terms of different parameters rather than indices?}

  We have followed your suggestion; \textbf{see the last paragraph of Section 2.1}.

\item \emph{Page 5, ``that minimizes the compound risk.” $\rightarrow$ “that minimizes an estimate of the compound risk''?}

  Fixed, thanks.

\item \emph{Page 5, after equation 3: Maybe explicitly write that $\beta_S$, $\beta_I$ are scalars (i.e., $\beta_S, \beta_I \in \mathbb{R}$)? Has $\bs{I}$ been explicitly defined as the identity matrix? }

  We have now written that $\beta_S$ and $\beta_I$ are scalars are have identified $\bs{I}$ as the $p \times p$ identity matrix; \textbf{see after equation (3)}.

\item \emph{Page 9: ``In this case, the optimal separable rule for estimating $\sigma_{jk}$ borrows from all of the $\sigma_j$ and $r_{jk}$'' $\rightarrow$ Perhaps use different indices when referring to the first $\sigma_{jk}$ and when referring to all parameters? }

  We have now reworded this as ``In this case, the optimal separable rule for estimating the $jk$th entry of $\bs{\Sigma}$ borrows from all of the $\sigma_j$ and $r_{jk}$''; \textbf{see after (11)}.

\item \emph{Page 10: ``We propose to calculate estimate'' $\rightarrow$ ``we propose to estimate''?}

  Fixed, thanks.

\item \emph{Page 18: ``The data are from the NCBI'' $\rightarrow$ ``The data are available from the NCBI''?}

  Fixed, thanks.

\item \emph{Page 21: ``These imposes significant'' $\rightarrow$ ``impose'' }

  Fixed, thanks.

\end{enumerate}

\bibliographystyle{abbrvnat}
\bibliography{biblio}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2022-03-08 msg.tex
edited to satisfy reviewer 2's comments

The article is organized as follows. In Section \ref{sec:method}, we briefly review compound decision theory and then introduce our proposed estimator. In Section \ref{numerical results} we illustrate its performance in simulations and a gene coexpression network inference problem. Section \ref{sec:discussion} concludes with a discussion, and proofs can be found in the Supporting Information.

@@

Given $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$, our goal is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$. A common measure of estimation performance is the scaled squared Frobenius risk
\begin{equation}
\label{frobenius risk}
R(\bs{\Sigma}, \bs{\delta}) = \frac{1}{p^2} \sum_{j,k=1}^p \mathbb{E}[\{\delta_{jk}(\bs{X})-\sigma_{jk}\}^2],
\end{equation}
where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate.

@@

Covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} is thus a compound decision problem where the goal is to simultaneously estimate the components of
\begin{equation}
  \label{eq:sigma}
  \bs{\sigma} = (\sigma_{11}, \ldots, \sigma_{1p}, \ldots, \sigma_{p1}, \ldots,\sigma_{pp})^\top
\end{equation}
under average squared error loss.  One important difference between estimating a covariance matrix versus a vector is that the former should have additional structure, and in particular should be at least positive semidefinite. Notably, however, this structure is not incentived by the Frobenius risk \eqref{frobenius risk}. As a result, there can exist estimators of $\bs{\Sigma}$ that achieve low values of \eqref{frobenius risk} but which are not positive-definite. To resolve this issue, we can project any estimate into the space of positive-definite matrices; see Section \ref{posdef}.

The value of this formulation is that it allows us to apply ideas from the compound decision literature to covariance matrix estimation.  A key property of compound decision problems is that while information about index $i$ seems like it should offer no help for estimating a parameter with a different index $j$, in fact borrowing information across all indices is superior to considering each index alone. An example of this phenomenon is the James-Stein shrinkage estimator \citep{james1961estimation}% , which estimates $\bs{\theta} = E \bs{Y}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$. \citet{james1961estimation} showed that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$
, and a long line of subsequent work has led to much more sophisticated and accurate procedures \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}. In Section \ref{sec:method} we apply these ideas to covariance matrix estimation,

@@

As a further illustration of the potential of our new formulation, we can show that the linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be essentially recovered as the solution to a compound decision problem. A standard approach in this literature is to first define a class of decision rules for estimating each component and then identify the member of that class that minimizes the compound risk % . For example, in the Gaussian sequence problem the Efron-Morris estimator \citep{efron1973stein} can be viewed as the optimal linear decision rule 
\citep{fourdrinier2018shrinkage, stigler19901988}. Following this recipe, we consider the class of linear decision rules where $\sigma_{jk}$ will be estimated by
\begin{equation}
  \label{linear class}
  \delta_{jk}(\bs{X}) = \beta_S s_{jk} + \beta_I u_{jk}
\end{equation}
for some parameters $\beta_S$ and $\beta_I$, where $s_{jk}$ is the $jk$th entry of $\bs{S}$ and $u_{jk}$ is the $jk$th entry of $\bs{I}$. Ideally we would like to estimate the optimal $\beta_S$ and $\beta_I$ by choosing the values that minimize the Frobenius risk \eqref{frobenius risk}. Though we cannot directly calculate \eqref{frobenius risk} because it involves the unknown $\sigma_{jk}$, the following proposition shows that we can construct a good estimate of this risk for every fixed value of $\beta_S$ and $\beta_I$.

@@

Proposition \ref{prop:bayes risk} illustrates the key property of compound decision problems: that borrowing information across indices can help estimate a single parameter. In this case, the optimal separable rule for estimating $\sigma_{jk}$ borrows from all of the $\sigma_j$ and $r_{jk}$. This rule also has an interesting Bayesian interpretation, as $t^\star_{od}$ and $t^\star_d$ are formally equivalent to posterior expectations when the $(\sigma_j, \sigma_k, r_{jk})$ are independently and identically distributed according to the prior $G_{od}$. However, the result in Proposition \ref{prop:bayes risk} holds even when the parameters are not random. Proposition \ref{prop:bayes risk} is a direct extension of the fundamental theorem of compound decisions \citep{robbins1951asymptotically, zhang2003compound}.

The optimal $\bs{\delta}^\star$ \eqref{eq:opt} is an oracle estimator and cannot be calculated in practice. A common approach in the compound decision literature is to use empirical Bayes ideas to estimate the oracle optimal separable decision rule \citep{robbins1955empirical, zhang2003compound, brown2009nonparametric, jiang2009general, efron2014two, efron2019bayes}. We propose to estimate $\bs{\delta}^\star$ using what \citet{efron2014two} calls $g$-modeling, where we proceed as if the $(\sigma_j, \sigma_k, r_{jk})$ were truly random and then estimate the prior $G_{od}$ from the data. Specifically, under the working assumption that the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are independent for different $(j, k)$, we propose to calculate estimate $G_{od}$ and $G_d$ \eqref{eq:G} using nonparametric maximum likelihood \citep{kiefer1956consistency}:
\begin{equation}
  \label{Gnd hat}
  \hat{G}_{od} = \argmax_{G_{od} \in \mathcal{G}_{od}}
  \prod_{1\leq k < j \leq p} \int f_{2n}(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid a, b, \gamma) dG_{od}(a, b, \gamma)
  \prod_{j = 1}^p \int f_{1n} (\bs{X}_{\cdot j} \mid a) dG_d(a),
\end{equation}
where $\mathcal{G}_{od}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$ and $G_d$ is determined by $G_{od}$ as indicated in \eqref{eq:G}. Of course, the $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$ are not independent across different $(j, k)$ unless the true covariance $\bs{\Sigma}$ is diagonal, so $\hat{G}_{od}$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}. Using $\hat{G}_{od}$ and $\hat{G}_d$, we propose to estimate the vectorized $\bs{\Sigma}$ using
\begin{equation}
  \label{proposed}
  \hat{\bs{\delta}}(\bs{X})
  =
  (
  \hat{t}_d(\bs{X}_{\cdot 1}), \ldots, \hat{t}_{od}(\bs{X}_{\cdot 1}, \bs{X}_{\cdot p}), \ldots, \hat{t}_{od}(\bs{X}_{\cdot p}, \bs{X}_{\cdot 1}), \ldots, \hat{t}_d(\bs{X}_{\cdot p})
  )
\end{equation}
where $\hat{t}_{od}$ and $\hat{t}_d$ are obtained by plugging $\hat{G}_{od}$ and $\hat{G}_d$ \eqref{Gnd hat} into the expressions for $t_{od}^\star$ and $t_d^\star$ from \eqref{eq:tstar}.

@@

In addition to the above data-driven estimators, we also implemented two oracle estimators, which cannot be implemented in practice as they require the unknown $\bs{\Sigma}$: (a) OracNonlin: the optimal rotation-invariant covariance estimator, derived in \citet{ledoit2019quadratic}, of $\bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{U}=(\bs{u}_1\ldots\bs{u}_p)$ is the sample eigenvector matrix and $\bs{l} = (d_1, \ldots, d_p)$ is composed of oracle eigenvalues $d_i = \bs{u}_i^T\bs{\Sigma} \bs{u}_i$; and (b) OracMSG: we implemented our proposed estimator \eqref{proposed} except that the support points are generated by clustering the true parameters $(\sigma_j,\sigma_k,r_{jk})$ instead of their sample versions.

@@

To measure the accuracy of the estimators, we randomly split the samples into a training and testing set. We estimated the covariance matrix on the training set and measured accuracy using the Frobenius norm of the difference between the estimated matrix and the sample covariance matrix calculated from the test set. We set the training sample size equal to 10, which left five samples in the test sets for the amygdala and frontal cortex and only three samples in the test set for hypothalamus. We therefore dropped hypothalamus from the analysis because we felt that three samples was unsuitable for calculating a surrogate of the true covariance matrix. We repeated this process 100 times. Table \eqref{tab:tab1} reports the median errors and interquartile ranges across the replications. Many of the different methods had comparable performance, and the different brain regions favored different methods. Nevertheless, our proposed MSGCor had among the best performances in both regions.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.9\textwidth]{img/networks_A.png}
  \end{center}
\caption{Amygdala gene networks recovered by the different covariance matrix estimation methods.}
\label{network_A}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=0.9\textwidth]{img/networks_FC.png}
  \end{center}
\caption{Frontal cortex gene networks recovered by the different covariance matrix estimation methods.}
\label{network_FC}
\end{figure}

In addition to comparing the numerical accuracies, we also investigated whether our estimator gave qualitatively different gene networks compared to the other approaches. Figures \ref{network_A} and \ref{network_FC} illustrate the estimated amygdala and frontal cortex covariance matrices in network form, where each node represents a gene and each edge represents a non-zero covariance between the genes it connects. To avoid completely connected graphs, we sparsified the matrix estimates by thresholding the smaller entries of each matrix to zero. Since the adaptive thresholding method of \citet{cai2011adaptive} naturally produced a sparse estimated matrix, we thresholded the other matrix estimates to match the sparsity level of the \citet{cai2011adaptive} estimate. The network representations suggest that the resulting networks look fairly similar except for the ones produced by linear shrinkage and NERCOME. That our procedure produces networks that look similar to the unbiased sample covariance matrix networks is comforting.

@@

Providing theoretical guarantees for our estimator is difficult. In the standard mean vector estimation problem with $Y_i \sim N(\theta_i, 1)$, \citet{jiang2009general} showed that an empirical Bayes estimator based on a nonparametric maximum likelihood estimate of the prior on the $\theta_i$ can indeed asymptotically achieve the same risk as the oracle optimal separable estimator. However, this was in a simple model with a univariate prior distribution. \citet{saha2020nonparametric} extended these results to multivariate $\bs{Y}_i \sim N(\bs{\theta}_i, \bs{\Sigma}_i)$ with a multivariate prior on the $\bs{\theta}_i$, but assumed that the $\bs{Y}_i$ were independent. In contrast, our covariance matrix estimator is built from arbitrarily dependent $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$. These imposes significant theoretical difficulties that will require substantial work to address; we leave this for future research.

>>>>

The article is organized as follows. In Section \ref{sec:compound} we briefly review compound decision theory, and in Section \ref{sec:method} we introduce our proposed estimator. In Section \ref{numerical results} we illustrate its performance in simulations and Section \ref{gene analysis} contains an analysis of the gene coexpression network inference problem. Section \ref{sec:discussion} concludes with a discussion, and proofs can be found in the Supporting Information.

@@

Given $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$, our goal is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$. A common measure of estimation performance is the scaled squared Frobenius risk
\begin{equation}
\label{frobenius risk}
R(\bs{\Sigma}, \bs{\delta}) = \frac{1}{p^2} \sum_{j,k=1}^p \mathbb{E}[\{\delta_{jk}(\bs{X})-\sigma_{jk}\}^2],
\end{equation}
where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate and $\sigma_{jj} = \sigma^2_j$ is the variance of the $j$th feature.

@@

Covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} is thus a compound decision problem where the goal is to simultaneously estimate the components of
\begin{equation}
  \label{eq:sigma}
  \bs{\sigma} = (\sigma_{11}, \ldots, \sigma_{1p}, \ldots, \sigma_{p1}, \ldots,\sigma_{pp})^\top
\end{equation}
under average squared error loss.  One important difference between estimating a covariance matrix versus a vector is that the former should have additional structure, and in particular should be at least positive semidefinite. Notably, however, this structure is not incentivized by the Frobenius risk \eqref{frobenius risk}. As a result, there can exist estimators of $\bs{\Sigma}$ that achieve low values of \eqref{frobenius risk} but which are not positive-definite. To resolve this issue, we can project any estimate into the space of positive-definite matrices; see Section \ref{posdef}.

The value of this formulation is that it allows us to apply ideas from the compound decision literature to covariance matrix estimation.  A key property of compound decision problems is that while information about a parameter $\theta_i$ seems like it should offer no help for estimating any other $\theta_j$, in fact borrowing information across all $\theta_i$ is superior to considering each parameter separately. An example of this phenomenon is the James-Stein shrinkage estimator \citep{james1961estimation}% , which estimates $\bs{\theta} = E \bs{Y}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$. \citet{james1961estimation} showed that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$
, and a long line of subsequent work has led to much more sophisticated and accurate procedures \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}. In Section \ref{sec:method} we apply these ideas to covariance matrix estimation.

@@

As a further illustration of the potential of our new formulation, we can show that the linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be essentially recovered as the solution to a compound decision problem. A standard approach in this literature is to first define a class of decision rules for estimating each component and then identify the member of that class that minimizes an estimate of the compound risk % . For example, in the Gaussian sequence problem the Efron-Morris estimator \citep{efron1973stein} can be viewed as the optimal linear decision rule 
\citep{fourdrinier2018shrinkage, stigler19901988}. Following this recipe, we consider the class of linear decision rules where $\sigma_{jk}$ will be estimated by
\begin{equation}
  \label{linear class}
  \delta_{jk}(\bs{X}) = \beta_S s_{jk} + \beta_I u_{jk}
\end{equation}
for some scalar parameters $\beta_S$ and $\beta_I$, where $s_{jk}$ is the $jk$th entry of $\bs{S}$ and $u_{jk}$ is the $jk$th entry of the $p \times p$ identity matrix $\bs{I}$. Ideally we would like to estimate the optimal $\beta_S$ and $\beta_I$ by choosing the values that minimize the Frobenius risk \eqref{frobenius risk}. Though we cannot directly calculate \eqref{frobenius risk} because it involves the unknown $\sigma_{jk}$, the following proposition shows that we can construct a good estimate of this risk for every fixed value of $\beta_S$ and $\beta_I$.

@@

Proposition \ref{prop:bayes risk} illustrates the key property of compound decision problems: that borrowing information across indices can help estimate a single parameter. In this case, the optimal separable rule for estimating the $jk$th entry of $\bs{\Sigma}$ borrows from all of the $\sigma_j$ and $r_{jk}$. This rule also has an interesting Bayesian interpretation, as $t^\star_{od}$ and $t^\star_d$ are formally equivalent to posterior expectations when the $(\sigma_j, \sigma_k, r_{jk})$ are independently and identically distributed according to the prior $G_{od}$. However, the result in Proposition \ref{prop:bayes risk} holds even when the parameters are not random. Proposition \ref{prop:bayes risk} is a direct extension of the fundamental theorem of compound decisions \citep{robbins1951asymptotically, zhang2003compound}.

The optimal $\bs{\delta}^\star$ \eqref{eq:opt} is an oracle estimator and cannot be calculated in practice. A common approach in the compound decision literature is to use empirical Bayes ideas to estimate the oracle optimal separable decision rule \citep{robbins1955empirical, zhang2003compound, brown2009nonparametric, jiang2009general, efron2014two, efron2019bayes}. We propose to estimate $\bs{\delta}^\star$ using what \citet{efron2014two} calls $g$-modeling, where we proceed as if the $(\sigma_j, \sigma_k, r_{jk})$ were truly random and then estimate the prior $G_{od}$ from the data. Specifically, under the working assumption that the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are independent for different $(j, k)$, we propose to estimate $G_{od}$ and $G_d$ \eqref{eq:G} using nonparametric maximum likelihood \citep{kiefer1956consistency}:
\begin{equation}
  \label{Gnd hat}
  \hat{G}_{od} = \argmax_{G_{od} \in \mathcal{G}_{od}}
  \prod_{1\leq k < j \leq p} \int f_{2n}(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid a, b, \gamma) dG_{od}(a, b, \gamma)
  \prod_{j = 1}^p \int f_{1n} (\bs{X}_{\cdot j} \mid a) dG_d(a),
\end{equation}
where $\mathcal{G}_{od}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$ and $G_d$ is determined by $G_{od}$ as indicated in \eqref{eq:G}. Of course, the $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$ are not independent across different $(j, k)$ unless the true covariance $\bs{\Sigma}$ is diagonal, so $\hat{G}_{od}$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}. Using $\hat{G}_{od}$ and $\hat{G}_d$, we propose to estimate the vectorized $\bs{\Sigma}$ using
\begin{equation}
  \label{proposed}
  \hat{\bs{\delta}}(\bs{X})
  =
  (
  \hat{t}_d(\bs{X}_{\cdot 1}), \ldots, \hat{t}_{od}(\bs{X}_{\cdot 1}, \bs{X}_{\cdot p}), \ldots, \hat{t}_{od}(\bs{X}_{\cdot p}, \bs{X}_{\cdot 1}), \ldots, \hat{t}_d(\bs{X}_{\cdot p})
  )
\end{equation}
where $\hat{t}_{od}$ and $\hat{t}_d$ are obtained by plugging $\hat{G}_{od}$ and $\hat{G}_d$ \eqref{Gnd hat} into the expressions for $t_{od}^\star$ and $t_d^\star$ from \eqref{eq:tstar}.

@@

In addition to the above data-driven estimators, we also implemented two oracle estimators, which cannot be implemented in practice as they require the unknown $\bs{\Sigma}$: (a) OracNonlin: the optimal rotation-invariant covariance estimator, derived in \citet{ledoit2019quadratic}, of $\bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{U}=(\bs{u}_1\ldots\bs{u}_p)$ is the sample eigenvector matrix and $\bs{l} = (d_1, \ldots, d_p)$ is composed of oracle eigenvalues $d_i = \bs{u}_i^T\bs{\Sigma} \bs{u}_i$; and (b) OracMSG: we implemented our proposed estimator \eqref{proposed} except that the support points are generated by clustering the true parameters $(\sigma_j,\sigma_k,r_{jk})$ instead of their sample versions. We performed clustering even though the true parameters are all known in order to reduce the number of support points to ease the computational burden.

@@

To measure the accuracy of the estimators, we randomly split the samples into a training and testing set. For the sake of sample size, we have ignored the timepoints at which the gene expressions were measured, following \citet{saul2017transcriptional}. We estimated the covariance matrix on the training set and measured accuracy using the Frobenius norm of the difference between the estimated matrix and the sample covariance matrix calculated from the test set. We set the training sample size equal to 10, which left five samples in the test sets for the amygdala and frontal cortex and only three samples in the test set for hypothalamus. We therefore dropped hypothalamus from the analysis because we felt that three samples was unsuitable for calculating a surrogate of the true covariance matrix. We repeated this process 100 times. Table \eqref{tab:tab1} reports the median errors and interquartile ranges across the replications. Many of the different methods had comparable performance, and the different brain regions favored different methods. Nevertheless, our proposed MSGCor had among the best performances in both regions.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.45\textheight]{img/networks_A_2.png}
  \end{center}
\caption{Amygdala gene networks recovered by the top performers in Table \ref{tab:tab1}. Line types indicate the method(s) that recovered each each edge.}
\label{network_A}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=0.45\textheight]{img/networks_FC_2.png}
  \end{center}
\caption{Frontal cortex gene networks recovered by the top performers in Table \ref{tab:tab1}. Line types indicate the method(s) that recovered each each edge.}
\label{network_FC}
\end{figure}

In addition to comparing the numerical accuracies, we also compared the gene networks produced by applying the various approaches to all samples. To avoid completely connected graphs, we sparsified the matrix estimates by thresholding the smaller entries of each matrix to zero. Since the adaptive thresholding method of \citet{cai2011adaptive} naturally produced a sparse estimated matrix, we thresholded the other matrix estimates to match the sparsity level of the \citet{cai2011adaptive} estimate. Figures \ref{network_A} and \ref{network_FC} illustrate the estimated amygdala and frontal cortex covariance matrix estimates from the best performers from Table \ref{tab:tab1} in network form, where each node represents a gene and each edge represents a non-zero covariance between the genes it connects. The figures show that our MSGCor approach gave qualitatively different gene networks. This suggests a useful strategy for estimating gene networks in practice: high-confidence edges can be identified as those edges recovered by multiple covariance matrix estimators.

@@

Providing theoretical guarantees for our estimator is difficult. In the standard mean vector estimation problem with $Y_i \sim N(\theta_i, 1)$, \citet{jiang2009general} showed that an empirical Bayes estimator based on a nonparametric maximum likelihood estimate of the prior on the $\theta_i$ can indeed asymptotically achieve the same risk as the oracle optimal separable estimator. However, this was in a simple model with a univariate prior distribution. \citet{saha2020nonparametric} extended these results to multivariate $\bs{Y}_i \sim N(\bs{\theta}_i, \bs{\Sigma}_i)$ with a multivariate prior on the $\bs{\theta}_i$, but assumed that the $\bs{Y}_i$ were independent. In contrast, our covariance matrix estimator is built from arbitrarily dependent $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$. These impose significant theoretical difficulties that will require substantial work to address; we leave this for future research.
